### 2024-08-17 paper.tex
work on section 8

\newcommand{\leftshift}{\mathbin{\rotatebox[origin=c]{0}{$\ll$}}}
\newcommand{\downshift}{\mathbin{\rotatebox[origin=c]{90}{$\ll$}}}
\newcommand{\rightshift}{\mathbin{\rotatebox[origin=c]{180}{$\ll$}}}
\newcommand{\upshift}{\mathbin{\rotatebox[origin=c]{270}{$\ll$}}}

@@

\section{Finding the irreducible components of \(U_0\)}
Let \(X\) be nilpotent of shape \(\lambda\).
Let 
\[U_0 = \{\]
Start fresh finding irreducible components of \(U_{w0}\).
\par To do this, we will use a known result about the irreducible components of the usual Springer fiber at an element \(J(\mu) \in \gl_{m + n}\).
(I should define this somewhere else, but \(J(\mu)\) just has the blocks in order of nonincreasing size.)
\begin{theorem}
    (Needs citation!)
    The irreducible components of the Springer fiber at \(J(\mu)\) are in bijection with the standard Young tableaus of shape \(\mu\).
    Further, the irreducible components are equidimensional, of dimension \(\sum_{i \neq j} \min(\mu_i, \mu_j) = \sum_i (i - 1) \mu_i\).
\end{theorem}
Let \(W_\mu\) denote the Springer fiber at \(J(\mu)\), and let \((W_{\mu, d})_d\) be the irreducible components.
We remark that the Springer fiber at \(PJ(\mu)P^{-1}\) is \(PW_\mu\), and its irreducible components are \((PW_{\mu, d})_d\).
We will use this result; to find the regular Springer fiber at any \(X \in \gl_{m + n}\), we just need to find \(P\) such that \(P^{-1}XP\) is in Jordan normal form; equivalently, we just need to find a Jordan basis for \(X\).
So, our strategy will be to find a subset of \(U_{wr}\) such that \(A_{e_{wr},b}\) has some fixed Jordan form \(\mu\), and the map \(b \mapsto P\) is algebraic.
Then we can break this subset into irreducible components by putting the Springer fiber at \(b\) in correspondence with one of the \(W_{\mu,d}\).
I think this explanation of intuition is important, so I should try to make it better.
\par So, with this as our motivation, we find a Jordan basis for the matrices \(A\) with \((A, U) \in U_{wr}\).

@@

\subsection{The general case}
Let \(X\) be nilpotent in Jordan form, and let \(a = \mathbbm{1}_{w,1}\), and let \(b\) be such that \(\forall j. \; b_{wj} = 0\).
We describe a Jordan basis for \(A_{X, a, b}\).

\par Let \(\mu\) be the result of removing the row \(w\) from \(\lambda\).
Then set \(q_i = \lambda_i - p_i\).  Let \(q_{i_0} = \lambda_{i_0} - r - p_{i_0}\).
Now, it will be more convenient to assume our indices are such that \(q_i\) is nonincreasing, so we will do that.
We list the chains in (roughly) order of decreasing length.
First, for each \(i\) such that \(q_i \geq n + r\), we take the chain of length \(p_i + q_i\) beginning with \(e_{i, p_i + q_i}\).
\par Now we handle \(e_{i_0, \lambda_{i_0}}\).
Let \(P = \max_{i : q_i < n + r} p_i\).
\[v = e_{i_0, \lambda_{i_0}} - (A^{r + n + P}e_{i_0, \lambda_{i_0}} >> r + n + P).\]
Note that \(A^{r + n} e_{i_0, \lambda_{i_0}} = b'\), so \(A^{r + n + P} e_{i_0, \lambda_{i_0}} = A^P b' = b' << P\).
Also note that if we shift \(b'\) left \(P\) times, we zero out all the rows \(i\) where \(q_i < n + r\).
This ensures that the operation of shifting \(b' << P\) right \(r + n + P\) times is invertible by applying \(A\), and thus shifting left, \(r + n + P\) times.
So, we take the chain of length \(r + n + P\) beginning with \(v\).
\par Now, it is clear that \(e_{i_0, p_{i_0} + q_{i_0} + i}\) is in the span of the chains we have listed so far for each \(i \geq 1\).
It is also clear that \(b' \leftshift k\) is in the span, for each \(k \geq 0\).
\par Now, we handle the \(i\) with \(q_i < n + r\).
\[v_i = A^{n + r - q_i} e_{i_0, \lambda_{i_0}} - \sum_{k = 1}^{p_i}b_{i,k}' e_{i,k + q_i}.\]
Note that \(A^{q_i} v_i\) is just \(b'\) with row \(i\) zeroed out.
And \(A^{n + r - q_i} e_{i_0, \lambda_{i_0}}\) is also in the span of chains already listed, so it suffices to show that the closure of \(\langle \sum_k b'_{ik} e_{i,k + q_i}\rangle \) under action by \(A\) contains \(\langle e_{ij}\rangle\).
And this just follows from the fact that \(b'_{i,p_i}\) is nonzero.

>>>>

\newcommand{\leftshift}{$\,\texttt{<<}\,$}
\newcommand{\downshift}{\mathbin{\rotatebox[origin=c]{90}{\leftshift}}}
\newcommand{\rightshift}{\mathbin{\rotatebox[origin=c]{180}{\leftshift}}}
\newcommand{\upshift}{\mathbin{\rotatebox[origin=c]{270}{\leftshift}}}

@@

\section{Finding the irreducible components of \(U_0\)}
Let \(X\) be nilpotent with Jordan basis \((e_{ij})_{1 \leq i \leq k, 1 \leq j \leq \lambda_i}\).
Let \(V = \{(A_{a,b}, U) : \forall i. \; A_{a,b} U_{i + 1} \subseteq U\}\).
Let
\[U_0 = \{(A_{0,b}, U) \in V\}.\]
We will find the irreducible components of \(U_0\).
We note that even finding the irreducible components of \(\{A_{0,0}, U \in V\}\) is a pretty hard problem.
Not ambitious enough to try to re-solve that problem, we will use its solution.
\par To find the irreducible components of \(U_0\), we will use a known result about the irreducible components of the usual Springer fiber at an element \(J(\mu) \in \gl_{m + n}\).
(I should define this somewhere else, but \(J(\mu)\) just has the blocks in order of nonincreasing size.)
\begin{theorem}
    (Needs citation!)
    The irreducible components of the Springer fiber at \(J(\mu)\) are in bijection with the standard Young tableaus of shape \(\mu\).
    Further, the irreducible components are equidimensional, of dimension \(\sum_{i \neq j} \min(\mu_i, \mu_j) = \sum_i (i - 1) \mu_i\).
\end{theorem}
Let \(W_\mu\) denote the Springer fiber at \(J(\mu)\), and let \((W_{\mu, d})_d\) be the irreducible components.
We remark that the Springer fiber at \(PJ(\mu)P^{-1}\) is \(PW_\mu\), and its irreducible components are \((PW_{\mu, d})_d\).
We will use this result; to find the regular Springer fiber at any \(X \in \gl_{m + n}\), we just need to find \(P\) such that \(P^{-1}XP\) is in Jordan normal form; equivalently, we just need to find a Jordan basis for \(X\).
\par So, our strategy will be to find a subset of \(U_0\) such that \(A_{0,b}\) has some fixed Jordan form \(\mu\), and we have an algebraic map \(b \mapsto P\) such that for \(A_{0,b}\) in the subset, \(PJ(\mu)P^{-1} = A_{0,b}\).
Then we can break this subset into irreducible components by putting the Springer fiber at \(b\) in correspondence with one of the \(W_{\mu,d}\).
I think this explanation of intuition is important, so I should try to make it better.
\par Finding the map \(b \mapsto P\) amounts to finding a Jordan basis for \(A_{0,b}\).
So, with this as our motivation, we find a Jordan basis for \(A_{0,b}\).

\par We write \(b_{ij}\) to denote the projection of \(b \in \mathbb{C}^m\) onto \(e_{ij}\).
Then set \(q_i = \lambda_i - p_i\).
Now, it will be more convenient to assume our indices are such that \(q_i\) is nonincreasing, so we will do that: from now on, we assume that \(q_i\) is nonincreasing.
\par We list the chains of the Jordan basis in (roughly) order of decreasing length.
First, for each \(i\) such that \(q_i \geq n\), we take the chain of length \(p_i + q_i\) beginning with \(e_{i, p_i + q_i}\).
\par Now we handle \(f_n\).
Let \(P = \max_{i : q_i < n} p_i\).
\[v = f_n - (A^{n + P}f_n \rightshift n + P).\]
Note that \(A^n f_n = b\), so \(A^{n + P} f_n = A^P b = b \leftshift P\).
By shifting \(b\) left \(P\) times, we zero out all the rows \(i\) where \(q_i < n\).
This ensures that the operation of shifting \(b' \leftshift P\) right \(n + P\) times is invertible by applying \(A\), and thus shifting left, \(n + P\) times.
So, we take the chain of length \(n + P\) beginning with \(v\).
\par Now, it is clear that each \(f_i\) is in the span of the chains we have listed so far.
It is also clear that \(b \leftshift k\) is in the span, for each \(k \geq 0\).
\par Now, we handle the \(i\) with \(q_i < n\).
\[v_i = A^{n - q_i} f_n - \sum_{k = 1}^{p_i}b_{i,k} e_{i,k + q_i}.\]
Note that \(A^{q_i} v_i\) is just \(b\) with row \(i\) zeroed out.
And \(A^{n - q_i} f_n\) is also in the span of chains already listed, so it suffices to show that the closure of \(\langle \sum_k b_{ik} e_{i,k + q_i}\rangle \) under action by \(A\) contains \(\langle e_{ij}\rangle\).
And this just follows from the fact that \(b_{i,p_i}\) is nonzero.
I list the purported basis here:
    \item TODO

@@
